Causal Discovery on Precarity and Depression

### 2025-01-03 draft_v1/draft_v1.tex
draft update_till_sec2(analysis)

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\usepackage{amssymb}
\usepackage{booktabs} % Add to your preamble for cleaner table lines
\usepackage{makecell} % Add to your preamble for multi-line cells
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\date{2024-12-30}
\begin{document}
\maketitle
\begin{abstract}
\noindent Understanding the causal mechanisms linking precarity factors
and depression is critical for developing effective interventions. This
study utilizes the HELIUS dataset to explore these relationships using
advanced causal discovery methods. By applying algorithms such as FCI,
and CCI, and combining traditional Gaussian CI tests with non-parametric
approaches like RCoT, we investigate how precarity factors---including
employment, social, financial, housing, and relational stress---affect
depression, both as a sum score and at the individual symptom level. Our
findings reveal that relational stress consistently emerges as a
potential causal factor for depression, while symptoms such as sleep
disturbances, guilt, and anhedonia are particularly sensitive to
external stressors, acting as potential early warning signals or
intervention points for prevention. Moreover, the results highlight
complexities in the data, including the influence of latent confounders
and the challenges of capturing cyclic relationships. Despite some
limitations, such as unresolved ambiguities in causal directions and
challenges with mixed data distributions, this study demonstrates the
utility of causal discovery tools in disentangling the intricate
interplay between social and mental health dynamics. By mapping these
causal structures into computational models, future research can
simulate intervention effects, providing actionable insights to mitigate
the impact of precarity on mental health. This study serves as a
foundational effort, offering both methodological advancements and
practical implications for addressing depression at a population level.
\end{abstract}

@@

There are numerous causal discovery algorithms available; however, in
this study, we focus on algorithms suited to the potential cyclic
relationships within our system. Specifically, we use FCI (Fast Causal
Inference) and CCI (Cyclic Causal Inference), both capable of accounting
for such cycles under certain conditions (Mooij \& Claassen, 2020;
Strobl, 2019). Additionally, we include the PC algorithm as a reference,
given its simplicity and prominence as one of the most widely known
causal discovery methods (Spirtes et al., 2001). For a more detailed
explanation of these algorithms, please refer to Park et al. (2024).

\begin{table}[ht]
\centering
\caption{Assumptions of causal discovery algorithms}
\begin{tabular}{lccccl}
\toprule
Algorithm & Acyclicity & Causal sufficiency & 
\makecell{Absence of \\ selection bias} & Output \\ 
\midrule
PC  & $\checkmark$ & $\checkmark$ & $\checkmark$ & CPDAG \\ 
FCI & $-^{a}$ & $\checkmark$ & $-^{a}$ & PAG \\ 
CCI & x & x & x & MAAG \\ 
\bottomrule
\end{tabular}
\caption*{\small{\textit{Note}. $^{a}$FCI can detect cycles in systems that lack selection bias and exhibit non-linear relationships (Mooij \& Claassen, 2020).}}
\end{table}

As shown in Table 1, the resulting graphs from FCI and CCI differ
slightly (\emph{PAG}: partial ancestral graph; \emph{MAAG}: maximal
almost ancestral graph) due to their reliance on different underlying
assumptions. Both encode information about causal relationships between
variables, where the presence of an edge indicates causal
\emph{ancestry}. Directed edges, \(A\) *→ \(B\), specify that \(B\) is
not an ancestor of \(A\) in every graph within the Markov equivalence
class, \(Equiv(G)\). \(A\)*---\(B\) represent cases where \(B\) is an
ancestor of \(A\) in every graph in \(Equiv(G)\). Circle endpoints,
\(A\)*-o\(B\), denote ambiguity in the ancestral relationship, meaning
that \(B\)'s ancestral status relative to \(A\) varies across graphs in
\(Equiv(G)\).

In contrast, the graph produced by the PC algorithm is a \emph{CPDAG}
(completed partially directed acyclic graph), where directed edges
(\(A\) → \(B\)) indicate that \(A\) is a direct cause (parent) of \(B\).
Unlike FCI and CCI, the CPDAG does not include circle symbols. Instead,
when the PC algorithm cannot determine the direction, it represents
uncertainty with bidirectional arrows. While PC serves as a useful
reference, its strict assumptions of acyclicity and the absence of
latent confounders limit its applicability in more complex settings.
Therefore, our primary focus remains on the results from FCI and CCI.
For completeness, all PC algorithm results are provided in the
\hyperref[sec-appendix]{Appendix}.

One practical challenge in applying these algorithms to the HELIUS
dataset is that the data does not follow a Gaussian distribution, and
the relationships among variables are unlikely to be strictly linear. To
account for this, we complement the commonly used Gaussian conditional
independence test (CI test), which relies on partial correlations, with
a non-parametric CI test based on kernel methods. However, kernel-based
non-parametric tests are computationally demanding, particularly with
large datasets like ours. To mitigate this issue, we employ Randomized
Conditional Independence Test (RCIT) and the Randomized conditional
Correlation Test (RCoT), which uses random Fourier features to
approximate the kernel methods, thereby significantly reducing the
computational cost (Strobl et al., 2019). For a more detailed
explanation of RCIT and RCoT, see Section~\ref{sec-rcot}.

@@

We analyze the causal structure using three approaches: examining the
relationship (1) between the five individual precarity factors and the
PHQ sum score, which represents overall depression severity; (2) between
individual symptoms and precarity factors; and (3) between individual
symptoms and overall precarity (the sum of the five precarity factors).
Using the PHQ sum score simplifies the analysis by reducing
dimensionality, offering computational efficiency and a broad,
interpretable perspective on the relationship between depression and
precarity factors. In contrast, including individual symptoms provide a
more detailed understanding, capturing the diverse ways symptoms respond
to different precarity factors. However, this approach introduces
methodological challenges, such as non-standard distributions and the
complexities of increased dimensionality. Finally, examining individual
symptoms in relation to the overall precarity score complements the
analysis by revealing how symptoms collectively connect to general
precarity. By integrating these three approaches, we balance simplicity
and granularity, ensuring robust findings.

To address sensitivity to parameter values and enhance the robustness of
our results, we implement multiple settings combined with bootstrapping.
For each of the 100 bootstrap samples, we estimate causal graphs and
retain only the edges and directions that exceed predefined thresholds.
The analyses are conducted under the following conditions:

@@

This setup yields 16 combinations (2 significance levels × 4 thresholds
× 2 CI tests), applied across three algorithms. Each combination is
repeated for 100 bootstrap samples, yielding a total of 1,600 graphs per
algorithm. For analyses involving individual symptom variables, we
simplify the setup by focusing on thresholds of 0.6 and 0.7 and reducing
the number of bootstrap samples to 30 to ease computational demands.

To summarize the results, we identify the most frequently occurring edge
endpoints across different experimental setups.\footnote{The detailed
  proportion of each edge endpoint occurrence is shown in
  Section~\ref{sec-propmatrix}.} This ensures that only stable and
consistent edges are retained, providing a clearer and more reliable
understanding of the relationships between precarity factors and
depression. See Figure~\ref{fig-workflow} for an overview of the
analysis workflow.

@@

\caption{\label{fig-sum}Resulting graphs of precarity factors and
depression sum score using FCI and CCI}

@@

\caption{\label{fig-sym}Resulting graphs of precarity factors and
individual depression symptoms using FCI and CCI}

@@

\caption{\label{fig-presum}Resulting graphs of precarity sum score and
individual depression symptoms using FCI and CCI}

@@

@@

@@

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-8-1.pdf}
\end{center}

@@

\begin{center}
\includegraphics[width=0.55\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-9-1.pdf}
\end{center}

\begin{center}
\includegraphics[width=0.7\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-10-1.pdf}
\end{center}

@@

@@

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-12-1.pdf}
\end{center}

@@

\begin{center}
\includegraphics[width=0.7\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-13-1.pdf}
\end{center}

@@

@@

\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-15-1.pdf}
\end{center}
\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-15-2.pdf}
\end{center}
\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-15-3.pdf}
\end{center}
\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-15-4.pdf}
\end{center}
\end{minipage}%

@@

\begin{center}
\includegraphics[width=0.6\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-16-1.pdf}
\end{center}

@@

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-17-1.pdf}
\end{center}

@@

@@

\caption{\label{fig-pc_sum}Resulting graphs of precarity factors and
depression sum score using PC}

@@

\caption{\label{fig-pc_sym}Resulting graphs of precarity factors and
individual depression symptoms using PC}

@@

\caption{\label{fig-pc_presum}Resulting graphs of precarity sum score
and individual depression symptoms using PC}

@@

\caption{\label{fig-sum-mat}Proportions of edge endpoint types for
graphs based on depression sum score}

@@

\caption{\label{fig-sym-mat}Proportions of edge endpoint types for
graphs based on individual depression symptoms}

@@

\caption{\label{fig-presum-mat}Proportions of edge endpoint types for
graphs based on individual depression symptoms with precarity sum score}

>>>>

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\usepackage{amssymb}
\usepackage{booktabs} % Add to your preamble for cleaner table lines
\usepackage{makecell} % Add to your preamble for multi-line cells
\usepackage{tikz}
\usetikzlibrary{arrows.meta, shapes.geometric}

%% Define TikZ arrow styles
\newcommand{\starstar}{% 
\begin{tikzpicture}[baseline=-3pt]
    \draw [{Rays[n=6]}-{Rays[n=6]}] (0,0) -- (0.55,0);
\end{tikzpicture}
}

\newcommand{\circstar}{% 
\begin{tikzpicture}
    \draw [{Circle[open]}-{Rays[n=6]}] (0,0) -- (0.55, 0);
\end{tikzpicture}
}

\newcommand{\starcirc}{% 
\begin{tikzpicture}
    \draw [{Rays[n=6]}-{Circle[open]}] (0,0) -- (0.55, 0);
\end{tikzpicture}
}

\newcommand{\stararrow}{% 
\begin{tikzpicture}
    \draw [{Rays[n=6]}-{Straight Barb[length=2.5pt]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\arrowstar}{% 
\begin{tikzpicture}
    \draw [{Straight Barb[length=2.5pt]}-{Rays[n=6]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\circarrow}{% 
\begin{tikzpicture}
    \draw [{Circle[open]}-{Straight Barb[length=2.5pt]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\tailcirc}{% 
\begin{tikzpicture}[baseline=-3pt] 
    \draw [-{Circle[open]}] (0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand{\circtail}{% 
\begin{tikzpicture}[baseline=-3pt] 
    \draw [{Circle[open]}-] (0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand{\circirc}{% 
\begin{tikzpicture}[baseline=-3pt] 
    \draw [{Circle[open]}-{Circle[open]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\tailstar}{% 
\begin{tikzpicture}[baseline=-3pt] 
    \draw [-{Rays[n=6]}] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\startail}{% 
\begin{tikzpicture}
    \draw [{Rays[n=6]}-] (0,0) -- (0.5, 0);
\end{tikzpicture}
}

\newcommand{\tailarrow}{% 
\begin{tikzpicture}
    \draw [-{Straight Barb[length=2.5pt]}](0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand{\arrowtail}{% 
\begin{tikzpicture}
    \draw [{Straight Barb[length=2.5pt]}-](0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand{\arrowarrow}{% 
\begin{tikzpicture}
    \draw [{Straight Barb[length=2.5pt]}-{Straight Barb[length=2.5pt]}](0,0) -- (0.4, 0);
\end{tikzpicture}
}

\newcommand\stackedarrows{%
        \mathrel{\vcenter{\mathsurround0pt
                \ialign{##\crcr
                \noalign{\nointerlineskip}$\arrowtail$\crcr
                \noalign{\nointerlineskip}$\tailarrow$\crcr
                }
        }}%
}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\date{2025-01-03}
\begin{document}
\maketitle
\begin{abstract}
\noindent Understanding the causal mechanisms linking precarity factors
and depression is critical for developing effective interventions. This
study utilizes the HELIUS dataset to explore these relationships using
advanced causal discovery methods. By applying algorithms such as FCI,
and CCI, and combining traditional Gaussian CI tests with non-parametric
approaches like RCoT, we investigate how precarity factors---including
employment, social, financial, housing, and relational stress---affect
depression, both as a sum score and at the individual symptom level. Our
findings reveal that relational stress consistently emerges as a
potential causal factor for depression, while symptoms such as sleep
disturbances, guilt, and anhedonia are particularly sensitive to
external stressors, acting as potential early warning signals or
intervention points for prevention. Moreover, the results highlight
complexities in the data, including the influence of latent confounders
and the challenges of capturing cyclic relationships. Despite some
limitations, such as unresolved ambiguities in causal directions and
challenges with mixed data distributions, this study demonstrates the
utility of causal discovery tools in disentangling the intricate
interplay between social and mental health dynamics. By mapping these
causal structures into computational models, future research can
simulate intervention effects, providing actionable insights to mitigate
the impact of precarity on mental health. This study serves as a
foundational effort, offering both methodological advancements and
practical implications for addressing depression at a population level.
\end{abstract}

@@

There are numerous causal discovery algorithms available; however, in
this study, we focus on algorithms suited to the potential cyclic
relationships within our system. Specifically, we use FCI (Fast Causal
Inference) and CCI (Cyclic Causal Inference), both capable of accounting
for such cycles under certain conditions (Mooij \& Claassen, 2020;
Strobl, 2019). Additionally, we include the PC algorithm as a reference,
given its simplicity and prominence as one of the most widely known
causal discovery methods (Spirtes et al., 2001).

\begin{table}[ht]
\centering
\small % Make the font size smaller
\caption{Assumptions of causal discovery algorithms}
\begin{tabular}{lccccc}
\toprule
Algorithm & Acyclicity & \makecell{Causal \\ sufficiency} & 
\makecell{Absence of \\ selection bias} & Linearity & Output \\ 
\midrule
PC  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & CPDAG \\ 
FCI & $-^{a}$ & $\checkmark$ & $-^{a}$ & $-^{a}$ & PAG \\ 
CCI & x & x & x & $\checkmark$ & \makecell{\footnotesize(partially oriented) \\ MAAG} \\ 
\bottomrule
\end{tabular}
\caption*{\footnotesize{\textit{Note}. $^{a}$The FCI algorithm, introduced by Spirtes (1995), is a constraint-based causal discovery method for DAGs that accounts for latent confounding and selection bias. Mooij \& Claassen (2020) later showed its applicability to cyclic causal discovery with latent confounding under general faithfulness and Markov conditions, assuming non-linear causal relationships.}}
\end{table}

As shown in Table 1, the resulting graphs from FCI and CCI differ
slightly (\emph{PAG}: partial ancestral graph; \emph{MAAG}: maximal
almost ancestral graph) due to their reliance on different underlying
assumptions. Despite these differences, both graphs are forms of
\emph{ancestral graphs}, designed to encode causal relationships between
variables, where the presence of an edge indicates causal
\emph{ancestry.} Directed edges, \(A \stararrow B\), indicate that \(B\)
is not an ancestor of \(A\) in every graph within the Markov equivalence
class, \(Equiv(G)\). The Markov equivalence class refers to the set of
graphs that encode the same conditional independence relationships, such
that the same \emph{d-separation} relations hold across all graphs in
the class (Spirtes et al., 2001). \(A \startail B\) indicates that \(B\)
is an ancestor of \(A\) across all graphs in \(Equiv(G)\). Circle
endpoints, \(A \starcirc B\), represent ambiguity in the ancestral
relationship, meaning \(B\)'s ancestral status relative to \(A\) varies
across graphs in \(Equiv(G)\).\footnote{\(*\) serves as a
  \textit{meta-symbol}, representing one of the three possible
  edge-endpoints. For instance, \(A \tailstar B\) can indicate any of
  the following edges: \(A\) \textemdash~\(B\), \(A\tailarrow B\), or
  \(A \tailcirc B\) (Park et al., 2024).} When \(A \arrowarrow B\),
neither \(A\) nor \(B\) is an ancestor of the other, indicating the
presence of a latent confounder between them.

The algorithms also differ in their approaches and assumptions for
detecting cycles. In CCI's MAAG, \(A\) \textemdash~\(B\) indicates that
\(A\) is an ancestor of \(B\) and \(B\) is an ancestor of \(A\),
referring to a cycle between \(A \stackedarrows B\). FCI, on the other
hand, identifies potential cycles more subtly: fully-connected nodes
with circle endpoints (\(\circirc\)) in FCI's PAG may suggest a cyclic
structure. FCI basically provides a sufficient condition to distinguish
nodes that are not part of a cycle, offering a more nuanced approach
(Mooij \& Claassen, 2020). Table 1 highlights further differences in the
assumptions underlying these algorithms. CCI operates under the
assumption of a linear system, while FCI, particularly for inferring
cyclic relationships, assumes a non-linear system without selection bias
and adheres to more general faithfulness and Markov conditions (i.e.,
\emph{\(\sigma\)-separation} and \emph{\(\sigma\)-faithfulness} setting)
(Forré \& Mooij, 2018). When the respective assumptions of each
algorithm are satisfied, their outcomes for cyclic relationships align
with those illustrated in Figure~\ref{fig-examplegraphs}. While an
exhaustive discussion of causal discovery concepts and
algorithm-specific details is beyond the scope of this paper, readers
seeking a deeper understanding are encouraged to consult Park et al.
(2024) for comprehensive insights into these methods and their
applications.

\begin{minipage}{0.33\linewidth}

\includegraphics[width=0.6\textwidth,height=\textheight]{img/original_DCG.png}

\subcaption{\label{fig-examplegraphs-1}Example graph \(\mathcal{G}\)}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\includegraphics[width=0.6\textwidth,height=\textheight]{img/FCI_graph.png}

\subcaption{\label{fig-examplegraphs-2}Corresponding MAAG}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\includegraphics[width=0.6\textwidth,height=\textheight]{img/CCI_graph.png}

\subcaption{\label{fig-examplegraphs-3}Corrsponding PAG}

\caption{\label{fig-examplegraphs}Example graph \(\mathcal{G}\)
featuring a cycle and the corresponding MAAG from CCI and PAG from FCI.}

The graph produced by the PC algorithm is a \emph{CPDAG} (completed
partially directed acyclic graph), where directed edges (\(A\) → \(B\))
indicate that \(A\) is a direct cause (parent) of \(B\). Unlike FCI and
CCI, the CPDAG does not include circle symbols. Instead, when the PC
algorithm cannot determine the direction of causality, it represents
this uncertainty with bidirectional arrows. While the PC algorithm
serves as a useful reference, its strict assumptions of acyclicity and
the absence of latent confounders restrict its applicability in more
complex scenarios. Therefore, our primary focus remains on the results
from FCI and CCI, with all PC algorithm results provided in the
\hyperref[sec-appendix]{Appendix} for completeness.

One practical challenge in applying these algorithms to the HELIUS
dataset is that the data does not follow a Gaussian distribution, and
the relationships among variables are unlikely to be strictly linear. To
account for this, we supplement the commonly used Gaussian conditional
independence (CI) test, which relies on partial correlations, with a
non-parametric CI test based on reproducing kernel methods (Zhang et
al., 2012). While kernel-based conditional independence tests (KCIT) are
effective, they are computationally intensive for large datasets like
HELIUS, as their quadratic scaling with sample size is primarily driven
by the inversion of large kernel matrices. To mitigate this issue, we
use the Randomized Conditional Correlation Test (RCoT), which
approximate kernel methods using random Fourier features. This approach
reduces computation to linear scaling with sample size, significantly
lowering computational costs (Strobl et al., 2019). For a more detailed
explanation of RCoT, refer to Section~\ref{sec-rcot}.

@@

We analyze the causal structure using three approaches: examining the
relationship (1) between the five individual precarity factors and the
PHQ sum score, which represents overall depression severity; (2) between
individual symptoms and precarity factors; and (3) between individual
symptoms and overall precarity (the sum of the five precarity factors).
Using the PHQ sum score simplifies the analysis by reducing
dimensionality, offering computational efficiency and a broad,
interpretable perspective on the relationship between depression and
precarity factors. In contrast, analyzing individual symptoms provides a
more detailed understanding by capturing the diverse ways symptoms
respond to different precarity factors. However, this detailed approach
introduces methodological challenges, including non-standard
distributions of individual symptom variables and the complexities
associated with increased dimensionality. Finally, examining individual
symptoms in relation to the overall precarity score complements the
analysis by revealing how symptoms collectively connect to general
precarity. Aggregating precarity factors may uncover relationships that
could be overlooked in analyses of individual factors, particularly when
a symptom has associations across multiple factors that are not strong
enough to emerge in individual analyses. By integrating these three
approaches, we achieve a balance between capturing detailed granularity
and gaining a broader understanding of the relationship between
precarity and depression.

Furthermore, to address sensitivity to parameter values and enhance the
robustness of our results, we implement multiple settings combined with
bootstrapping. For each of the 100 bootstrap samples, we estimate causal
graphs and retain only the edges and directions that exceed predefined
thresholds. The analyses are conducted under the following conditions:

@@

This setup yields 16 combinations (2 significance levels × 4 thresholds
× 2 CI tests), applied across three algorithms. Each combination is
repeated for 100 bootstrap samples, yielding a total of 1,600 graphs per
algorithm. For analyses involving individual symptom variables, we
simplify the setup by focusing on thresholds of 0.6 and 0.7 and reducing
the number of bootstrap samples to 30, easing computational demands. To
further enhance efficiency, we fix the skeleton of edges among symptom
variables based on the common structure estimated across all algorithms.
This approach minimizes spurious edges in the symptom network and
significantly reduces the computational time needed for skeleton
estimation. This fixed structure also aligns well with commonly reported
skeleton structures in the literature, ensuring consistency with
existing findings (\textbf{cite the comp\_model paper}).\footnote{We
  fixed the skeleton of the symptom network using a common structure
  derived from the PC, FCI, and CCI algorithms with an alpha level of
  0.001. This approach preserved the complexity of symptom interactions,
  enabling unrestricted estimation of causal directions both among
  symptoms (\emph{symptom --- symptom} interactions) and between
  symptoms and precarity factors.} Lastly, to summarize the results, we
identify the most frequently occurring edge endpoints across different
experimental setups.\footnote{The detailed proportion of each edge
  endpoint occurrence is provided in Section~\ref{sec-propmatrix}.} This
ensures that only stable and consistent edges are retained, providing a
clearer and more reliable understanding of the relationships between
precarity factors and depression. See Figure~\ref{fig-workflow} for an
overview of the analysis workflow.

@@

\caption{\label{fig-sum}Resulting graphs of precarity factors and
depression sum score using FCI and CCI.}

@@

\caption{\label{fig-sym}Resulting graphs of precarity factors and
individual depression symptoms using FCI and CCI.}

@@

\caption{\label{fig-presum}Resulting graphs of precarity sum score and
individual depression symptoms using FCI and CCI.}

@@

\bibitem[\citeproctext]{ref-forre2018constraint}
Forré, P., \& Mooij, J. M. (2018). Constraint-based causal discovery for
non-linear structural causal models with cycles and latent confounders.
\emph{arXiv Preprint arXiv:1807.03024}.

@@

\bibitem[\citeproctext]{ref-spirtes_causal_1995}
Spirtes, P., Meek, C., \& Richardson, T. (1995). Causal inference in the
presence of latent variables and selection bias. \emph{Proceedings of
the {Eleventh} Conference on {Uncertainty} in Artificial Intelligence},
499--506.

@@

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-9-1.pdf}
\end{center}

@@

\begin{center}
\includegraphics[width=0.55\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-10-1.pdf}
\end{center}

@@

\begin{center}
\includegraphics[width=0.7\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-12-1.pdf}
\end{center}

@@

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-13-1.pdf}
\end{center}

@@

@@

\begin{center}
\includegraphics[width=0.7\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-15-1.pdf}
\end{center}

@@

\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-16-1.pdf}
\end{center}
\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-16-2.pdf}
\end{center}
\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-16-3.pdf}
\end{center}
\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-16-4.pdf}
\end{center}
\end{minipage}%

@@

\begin{center}
\includegraphics[width=0.6\textwidth,height=\textheight]{draft_v1_files/figure-pdf/unnamed-chunk-17-1.pdf}
\end{center}

@@

@@

\begin{center}
\includegraphics{draft_v1_files/figure-pdf/unnamed-chunk-19-1.pdf}
\end{center}

@@

\caption{\label{fig-pc_sum}Resulting graphs of precarity factors and
depression sum score using PC.}

@@

\caption{\label{fig-pc_sym}Resulting graphs of precarity factors and
individual depression symptoms using PC.}

@@

\caption{\label{fig-pc_presum}Resulting graphs of precarity sum score
and individual depression symptoms using PC.}

@@

\caption{\label{fig-sum-mat}Proportions of edge endpoint types for
graphs based on depression sum score.}

@@

\caption{\label{fig-sym-mat}Proportions of edge endpoint types for
graphs based on individual depression symptoms.}

@@

\caption{\label{fig-presum-mat}Proportions of edge endpoint types for
graphs based on individual depression symptoms with precarity sum
score.}
